---
title: "Doing Data Science Case Study 1"
author: "Jack Korbitz"
date: "2024-10-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Case Study 1

## Main Points:

|  |
|:-----------------------------------------------------------------------|
| 1\. Find the top three factors that contribute to Attrition |
| 2\. The business is also interested in learning about any job role specific trends that may exist in the data set |
| 3\. Provide any other interesting trends and observations from your analysis |
| 4\. Build a model to predict attrition |

## Load Packages
```{r}
library(tidyverse)
library(plotly)
library(e1071) # naiveBayes
library(caret) #cm tables
library(ggthemes)
library(scales)
```

## Load Data
```{r}
# load data
Attrition = read.csv("/Users/jackkorbitz/School/MSDS_6306/Unit 8/CaseStudy1-data.csv",header = TRUE)

# Check data structure
str(Attrition)

# Check data summary
summary(Attrition)

# Check for missing values
sum(is.na(Attrition))
```

## Explore different reasons for Attrition
```{r}
# Department Highest %: 21.61 @ 3 variables (Good Variable)
table(Attrition$Department,Attrition$Attrition)
Department_atr = table(Attrition$Department,Attrition$Attrition)
Department_atr_dist = prop.table(Department_atr, 1) *100
Department_atr_dist

# Age banded distribution: high for younger ages, not the best variable to use
Attrition$Age_Band <- cut(Attrition$Age, 
                         breaks = c(0, 29, 49,  60), 
                         labels = c("18-29", "30-49", "50-60"))

table(Attrition$Age_Band,Attrition$Attrition)
Age_atr = table(Attrition$Age_Band,Attrition$Attrition)
Age_atr_dist = prop.table(Age_atr, 1) *100
Age_atr_dist

# BusinessTravel Highest %: 22.15190 @ 3 variables
table(Attrition$BusinessTravel,Attrition$Attrition)
BusinessTravel_atr = table(Attrition$BusinessTravel,Attrition$Attrition)
BusinessTravel_atr_dist = prop.table(BusinessTravel_atr, 1) *100
BusinessTravel_atr_dist

# DistanceFromHome Highest %: 23.62205 @ 3 variables
Attrition$Distance_Band = cut(Attrition$DistanceFromHome, 
                         breaks = c(1, 10, 20, 30), 
                         labels = c("1-10", "11-20", "21-30"))
table(Attrition$Distance_Band,Attrition$Attrition)
DistanceFromHome_atr = table(Attrition$Distance_Band,Attrition$Attrition)
DistanceFromHome_atr_dist = prop.table(DistanceFromHome_atr, 1) *100
DistanceFromHome_atr_dist

# Education %: 18.36735 @ 5 variables (less education = higher turn)
table(Attrition$Education,Attrition$Attrition)
Education_atr = table(Attrition$Education,Attrition$Attrition)
Education_atr_dist = prop.table(Education_atr, 1) *100
Education_atr_dist

# EducationField %: 26.66667% @ 6 variable (Higher on lower counts)
table(Attrition$EducationField,Attrition$Attrition)
EducationField_atr = table(Attrition$EducationField,Attrition$Attrition)
EducationField_atr_dist = prop.table(EducationField_atr, 1) *100
EducationField_atr_dist

# EnvironmentSatisfaction %: 24.41860% @ 4 variables (highest at lowest and highest satisfaction, maybe needs a second variable)
table(Attrition$EnvironmentSatisfaction,Attrition$Attrition)
EnvironmentSatisfaction_atr = table(Attrition$EnvironmentSatisfaction,Attrition$Attrition)
EnvironmentSatisfaction_atr_dist = prop.table(EnvironmentSatisfaction_atr, 1) *100
EnvironmentSatisfaction_atr_dist

# Gender %: 16.86047% @ 2 variables (not useful by itself stand alone)
table(Attrition$Gender,Attrition$Attrition)
Gender_atr = table(Attrition$Gender,Attrition$Attrition)
Gender_atr_dist = prop.table(Gender_atr, 1) *100
Gender_atr_dist

# JobInvolvement %: 46.808511% @ 4 variables (Smaller samle but generally right)
table(Attrition$JobInvolvement,Attrition$Attrition)
JobInvolvement_atr = table(Attrition$JobInvolvement,Attrition$Attrition)
JobInvolvement_atr_dist = prop.table(JobInvolvement_atr, 1) *100
JobInvolvement_atr_dist

# JobLevel %: 26.139818% at large sample(good variable)
table(Attrition$JobLevel,Attrition$Attrition)
JobLevel_atr = table(Attrition$JobLevel,Attrition$Attrition)
JobLevel_atr_dist = prop.table(JobLevel_atr, 1) *100
JobLevel_atr_dist

# JobRole %: 45.283019% @ Sales rep (Low Job level Sales Dept,  identified)
table(Attrition$JobRole,Attrition$Attrition)
JobRole_atr = table(Attrition$JobRole,Attrition$Attrition)
JobRole_atr_dist = prop.table(JobRole_atr, 1) *100
JobRole_atr_dist

# We can see that department plays a major role, especially in sales. HR only has Attrition for level 1
p = Attrition |>
  group_by(JobLevel, Attrition, Department) |>
  summarise(count = n(), .groups = 'drop') |>
  ggplot(aes(x = factor(JobLevel), y = count, fill = Attrition)) +
  geom_bar(stat = "identity", position = "fill") +  
  scale_y_continuous(labels = scales::percent_format()) + 
  labs(title = "Attrition by Job Level (Proportion) by Department", x = "Job Level", y = "Proportion") +
  theme_minimal() +
  facet_wrap(~ Department)  
  ggplotly(p)
  
# JobSatisfaction %: 21.22905% at low satisfaction (Makes sense but doesn't really surprise us)
table(Attrition$JobSatisfaction,Attrition$Attrition)
JobSatisfaction_atr = table(Attrition$JobSatisfaction,Attrition$Attrition)
JobSatisfaction_atr_dist = prop.table(JobSatisfaction_atr, 1) *100
JobSatisfaction_atr_dist

table(Attrition$JobSatisfaction,Attrition$Department)

# Marital Status %: 26.022305% on single (Makes sense, probally not paying for as much and ease to move around country/ jobs)
table(Attrition$MaritalStatus,Attrition$Attrition)
MaritalStatus_atr = table(Attrition$MaritalStatus,Attrition$Attrition)
MaritalStatus_atr_dist = prop.table(MaritalStatus_atr, 1) *100
MaritalStatus_atr_dist

# NumCompaniesWorked %: 25.581395% lots of high turns with 1 other company or more than 5 (5-9)
table(Attrition$NumCompaniesWorked,Attrition$Attrition)
NumCompaniesWorked_atr = table(Attrition$NumCompaniesWorked,Attrition$Attrition)
NumCompaniesWorked_atr_dist = prop.table(NumCompaniesWorked_atr, 1) *100
NumCompaniesWorked_atr_dist

# OverTime %: 31.746032% High chance of turn on yes (Good variable, need to check with other jobs)
table(Attrition$OverTime,Attrition$Attrition)
OverTime_atr = table(Attrition$OverTime,Attrition$Attrition)
OverTime_atr_dist = prop.table(OverTime_atr, 1) *100
OverTime_atr_dist

table(Attrition$OverTime, Attrition$JobLevel)

# PercentSalaryHike %: Not a great variable by itself
table(Attrition$PercentSalaryHike,Attrition$Attrition)
PercentSalaryHike_atr = table(Attrition$PercentSalaryHike,Attrition$Attrition)
PercentSalaryHike_atr_dist = prop.table(PercentSalaryHike_atr, 1) *100
PercentSalaryHike_atr_dist

# StockOptionLevel %: 25.857520% suprised by high option level leaving, maybe retiring?
table(Attrition$StockOptionLevel,Attrition$Attrition)
StockOptionLevel_atr = table(Attrition$StockOptionLevel,Attrition$Attrition)
StockOptionLevel_atr_dist = prop.table(StockOptionLevel_atr, 1) *100
StockOptionLevel_atr_dist

# RelationshipSatisfaction %: 20.11494 not great by itself
table(Attrition$RelationshipSatisfaction,Attrition$Attrition)
RelationshipSatisfaction_atr = table(Attrition$RelationshipSatisfaction,Attrition$Attrition)
RelationshipSatisfaction_atr_dist = prop.table(RelationshipSatisfaction_atr, 1) *100
RelationshipSatisfaction_atr_dist

# TotalWorkingYears %: 24.427481 in 1-5 yr band
Attrition$Work_Years_Band <- cut(Attrition$TotalWorkingYears, 
                         breaks = c(1, 5, 10, 15, 20, 25, 30, 35, 40), 
                         labels = c("1-5", "6-10", "11-15","16-20", "21-25", "26-30", "31-35", "36-40"))
table(Attrition$Work_Years_Band,Attrition$Attrition)
TotalWorkingYears_atr = table(Attrition$Work_Years_Band,Attrition$Attrition)
TotalWorkingYears_atr_dist = prop.table(TotalWorkingYears_atr, 1) *100
TotalWorkingYears_atr_dist

# TrainingTimesLastYear %: 26.666667 not best identifier by itself
table(Attrition$TrainingTimesLastYear,Attrition$Attrition)
TrainingTimesLastYear_atr = table(Attrition$TrainingTimesLastYear,Attrition$Attrition)
TrainingTimesLastYear_atr_dist = prop.table(TrainingTimesLastYear_atr, 1) *100
TrainingTimesLastYear_atr_dist

# WorkLifeBalance %: 35.41667 not best identifier by itself, low sample
table(Attrition$WorkLifeBalance,Attrition$Attrition)
WorkLifeBalance_atr = table(Attrition$WorkLifeBalance,Attrition$Attrition)
WorkLifeBalance_atr_dist = prop.table(WorkLifeBalance_atr, 1) *100
WorkLifeBalance_atr_dist

# YearsAtCompany %: decent variable
table(Attrition$YearsAtCompany,Attrition$Attrition)
YearsAtCompany_atr = table(Attrition$YearsAtCompany,Attrition$Attrition)
YearsAtCompany_atr_dist = prop.table(YearsAtCompany_atr, 1) *100
YearsAtCompany_atr_dist

# YearsInCurrentRole %: decent variable
table(Attrition$YearsInCurrentRole,Attrition$Attrition)
YearsInCurrentRole_atr = table(Attrition$YearsInCurrentRole,Attrition$Attrition)
YearsInCurrentRole_atr_dist = prop.table(YearsInCurrentRole_atr, 1) *100
YearsInCurrentRole_atr_dist

# YearsSinceLastPromotion %: decent variable
table(Attrition$YearsSinceLastPromotion,Attrition$Attrition)
YearsSinceLastPromotion_atr = table(Attrition$YearsSinceLastPromotion,Attrition$Attrition)
YearsSinceLastPromotion_atr_dist = prop.table(YearsSinceLastPromotion_atr, 1) *100
YearsSinceLastPromotion_atr_dist

# YearsWithCurrManager %: decent variable
table(Attrition$YearsWithCurrManager,Attrition$Attrition)
YearsWithCurrManager_atr = table(Attrition$YearsWithCurrManager,Attrition$Attrition)
YearsWithCurrManager_atr_dist = prop.table(YearsWithCurrManager_atr, 1) *100
YearsWithCurrManager_atr_dist

# MonthlyIncome %: decent variable
Attrition$MonthlyIncome_Band <- cut(Attrition$MonthlyIncome, 
                         breaks = c(0, 4999, 9999,  14999, 20000), 
                         labels = c("0-4999", "5000-9999", "10k-14.9k","15k-20k"))
table(Attrition$MonthlyIncome_Band,Attrition$Attrition)
MonthlyIncome_atr = table(Attrition$MonthlyIncome_Band,Attrition$Attrition)
MonthlyIncome_atr_dist = prop.table(MonthlyIncome_atr, 1) *100
MonthlyIncome_atr_dist
```

## Factor 1: Overtime

```{r}
# viz 1
summary_data = Attrition |>
  group_by(OverTime, Attrition) |>
  summarise(count = n(), .groups = 'drop') |>
  mutate(percent = count / sum(count))  
  ggplot(summary_data, aes(x = OverTime, y = count, fill = Attrition, 
                              text = paste("Count:", count))) +  
  geom_bar(stat = "identity", position = "fill") +  
  scale_y_continuous(labels = percent_format()) + 
  labs(title = "Attrition: By OverTime",
       x = "OverTime",
       y = "Proportion") +
  theme_solarized_2() +
  scale_fill_manual(values = c("Yes" = scales::alpha("#EB172B", 0.75), "No" = scales::alpha("#F3AF49", 0.75))) +  
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 20)),  
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10)
  ) +  
  geom_text(aes(label = count), position = position_fill(vjust = 0.5), color = "white", size = 4)  


  
  
# viz 2
summary_data2 = Attrition |>
  group_by(JobLevel, OverTime, Attrition) |>
  summarise(count = n(), .groups = 'drop') |>
  mutate(percent = count / sum(count))  
  ggplot(summary_data2, aes(x = OverTime, y = count, fill = Attrition, 
                              text = paste("Count:", count))) +  
  geom_bar(stat = "identity", position = "fill") +  
  scale_y_continuous(labels = percent_format()) + 
  labs(title = "Attrition: By OverTime & Job Level",
       x = "OverTime",
       y = "Proportion") +
  theme_solarized_2() +
  scale_fill_manual(values = c("Yes" = scales::alpha("#EB172B", 0.75), "No" = scales::alpha("#F3AF49", 0.75))) +  
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 20)),  
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10)
  ) +  
  geom_text(aes(label = count), position = position_fill(vjust = 0.5), color = "white", size = 4) +  
  facet_wrap(~ JobLevel)  

# viz 3
summary_data3 = Attrition |>
  group_by(JobLevel, Attrition, Department, OverTime) |>
  summarise(count = n(), .groups = 'drop') |>
  mutate(percent = count / sum(count))  
  ggplot(summary_data3, aes(x = factor(JobLevel), y = count, fill = Attrition, 
                              text = paste("Count:", count))) +  
  geom_bar(stat = "identity", position = "fill") +  
  scale_y_continuous(labels = percent_format()) + 
  labs(title = "Attrition: By OverTime, Job Level, and Department",
       x = "Job Level",
       y = "Proportion") +
  theme_solarized_2() +
  scale_fill_manual(values = c("Yes" = scales::alpha("#EB172B", 0.75), "No" = scales::alpha("#F3AF49", 0.75))) +  
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 20)),  
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10)
  ) +  
  facet_wrap(~  OverTime + Department) +  
  geom_text(aes(label = count), position = position_fill(vjust = 0.5), color = "white", size = 4)  
```

## Factor 2: Business Travel by Age Band
```{r}
# Viz 1
Attrition$Age_Band = cut(Attrition$Age, 
                         breaks = c(0, 29, 49,  60), 
                         labels = c("18-29", "30-49", "50-60"))
  summary_data = Attrition |>
  group_by(Age_Band, Attrition) |>
  summarise(count = n(), .groups = 'drop') |>
  mutate(percent = count / sum(count))  
  ggplot(summary_data, aes(x = Age_Band, y = count, fill = Attrition, 
                              text = paste("Count:", count))) + 
  geom_bar(stat = "identity", position = "fill") +  
  scale_y_continuous(labels = percent_format()) + 
  labs(title = "Attrition: By Age Band",
       x = "Age",
       y = "Proportion") +
  theme_solarized_2() +
  scale_fill_manual(values = c("Yes" = scales::alpha("#EB172B", 0.75), "No" = scales::alpha("#F3AF49", 0.75))) +  
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 20)),  
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10)
  ) +  
  geom_text(aes(label = count), position = position_fill(vjust = 0.5), color = "white", size = 4) 


# Viz 2
summary_data2 = Attrition |>
  group_by(Age_Band, BusinessTravel, Attrition) |>
  summarise(count = n(), .groups = 'drop') |>
  mutate(percent = count / sum(count))  
  ggplot(summary_data2, aes(x = Age_Band, y = count, fill = Attrition, 
                              text = paste("Count:", count))) +  
  geom_bar(stat = "identity", position = "fill") +  
  scale_y_continuous(labels = percent_format()) + 
  labs(title = "Attrition: By Business Travel and Age Band",
       x = "Age",
       y = "Proportion") +
  theme_solarized_2() +
  scale_fill_manual(values = c("Yes" = scales::alpha("#EB172B", 0.75), "No" = scales::alpha("#F3AF49", 0.75))) +  
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 20)),  
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10)
  ) +  
  geom_text(aes(label = count), position = position_fill(vjust = 0.5), color = "white", size = 4) +  #
  facet_wrap(~ BusinessTravel)  

# viz 3
summary_data3 = Attrition |>
  group_by(Age_Band, Attrition, BusinessTravel, Department) |>
  summarise(count = n(), .groups = 'drop') |>
  mutate(percent = count / sum(count))
  ggplot(summary_data3, aes(x = factor(Age_Band), y = count, fill = Attrition,
                              text = paste("Count:", count))) +
  geom_bar(stat = "identity", position = "fill") +
  scale_y_continuous(labels = percent_format()) +
  labs(title = "Attrition: By Age Band, Business Travel and Gender",
       x = "Age",
       y = "Proportion") +
  theme_solarized_2() +
  scale_fill_manual(values = c("Yes" = scales::alpha("#EB172B", 0.75), "No" = scales::alpha("#F3AF49", 0.75))) +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 20)),
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10)
  ) +
  facet_wrap(~ Department + BusinessTravel) +
  geom_text(aes(label = count), position = position_fill(vjust = 0.5), color = "white", size = 4)

```

## Factor 3:

```{r}
# Viz 1
Attrition$MonthlyIncome_Band <- cut(Attrition$MonthlyIncome, 
                         breaks = c(0, 4999, 9999,  14999, 20000), 
                         labels = c("< $5k", "$5k-$9.9k", "$10k-$14.9k","> $15k"))
  summary_data = Attrition |>
  group_by(MonthlyIncome_Band, Attrition) |>
  summarise(count = n(), .groups = 'drop') |>
  mutate(percent = count / sum(count))  
  ggplot(summary_data, aes(x = MonthlyIncome_Band, y = count, fill = Attrition, 
                              text = paste("Count:", count))) + 
  geom_bar(stat = "identity", position = "fill") +  
  scale_y_continuous(labels = percent_format()) + 
  labs(title = "Attrition: By Monthly Income",
       x = "Monthly Income",
       y = "Proportion") +
  theme_solarized_2() +
  scale_fill_manual(values = c("Yes" = scales::alpha("#EB172B", 0.75), "No" = scales::alpha("#F3AF49", 0.75))) +  
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 20)),  
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10)
  ) +  
  geom_text(aes(label = count), position = position_fill(vjust = 0.5), color = "white", size = 4) 
  
# Viz 2
Attrition$Distance_Band <- cut(Attrition$DistanceFromHome, 
                         breaks = c(0, 10, 20, 30), 
                         labels = c("1-10", "11-20", "21-30"))
summary_data2 = Attrition |>
  group_by(MonthlyIncome_Band, Attrition, Distance_Band) |>
  summarise(count = n(), .groups = 'drop') |>
  mutate(percent = count / sum(count))  
  ggplot(summary_data2, aes(x = Distance_Band, y = count, fill = Attrition, 
                              text = paste("Count:", count))) + 
  geom_bar(stat = "identity", position = "fill") +  
  scale_y_continuous(labels = percent_format()) + 
  labs(title = "Attrition: By Monthly Income & Distance from Home",
       x = "Distance",
       y = "Proportion") +
  theme_solarized_2() +
  scale_fill_manual(values = c("Yes" = scales::alpha("#EB172B", 0.75), "No" = scales::alpha("#F3AF49", 0.75))) +  
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 20)),  
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10)
  ) +  
  geom_text(aes(label = count), position = position_fill(vjust = 0.5), color = "white", size = 4) +
  facet_wrap(~ MonthlyIncome_Band)  
  
# Viz 3
Attrition$Age_Band = cut(Attrition$Age, 
                         breaks = c(0, 29, 49,  60), 
                         labels = c("18-29", "30-49", "50-60"))
summary_data2 = Attrition |>
  group_by(MonthlyIncome_Band, Attrition, Age_Band) |>
  summarise(count = n(), .groups = 'drop') |>
  mutate(percent = count / sum(count))  
  ggplot(summary_data2, aes(x = Age_Band, y = count, fill = Attrition, 
                              text = paste("Count:", count))) + 
  geom_bar(stat = "identity", position = "fill") +  
  scale_y_continuous(labels = percent_format()) + 
  labs(title = "Attrition: By Monthly Income & Age",
       x = "Age",
       y = "Proportion") +
  theme_solarized_2() +
  scale_fill_manual(values = c("Yes" = scales::alpha("#EB172B", 0.75), "No" = scales::alpha("#F3AF49", 0.75))) +  
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 20)),  
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10)
  ) +  
  geom_text(aes(label = count), position = position_fill(vjust = 0.5), color = "white", size = 4) +
  facet_wrap(~ MonthlyIncome_Band)  
```

## Attrition model: Naive Bayes tuned
```{r}
# Load data
Attrition = read.csv("/Users/jackkorbitz/School/MSDS_6306/Unit 8/CaseStudy1-data.csv", header = TRUE)

# Define the columns in the model (not shown the amount of combinations tried...)
# Excludes to not be used 1,3,10,11
cols_to_use = c(2  # Age
               ,4  # BusinessTravel
             # ,5  # DailyRate
             # ,6  # Department
               ,7  # DistanceFromHome
             # ,8  # Education
             # ,9  # EducationField
             # ,10 # EmployeeCount
             # ,11 # EmployeeNumber
               ,12 # EnvironmentSatisfaction
             # ,13 # Gender
             # ,14 # HourlyRate
               ,15 # JobInvolvement
               ,16 # JobLevel
               ,17 # JobRole
               ,18 # JobSatisfaction
               ,19 # MaritalStatus
               ,20 # MonthlyIncome
             # ,21 # MonthlyRate
               ,22 # NumCompaniesWorked
             # ,23 # Over18
               ,24 # OverTime
             # ,25 # PercentSalaryHike
             # ,26 # PerformanceRating
               ,27 # RelationshipSatisfaction
             # ,28 # StandardHours    
               ,29 # StockOptionLevel
               ,30 # TotalWorkingYears
             # ,31 # TrainingTimesLastYear
               ,32 # WorkLifeBalance
               ,33 # YearsAtCompany
               ,34 # YearsInCurrentRole
             # ,35 # YearsSinceLastPromotion
               ,36 # YearsWithCurrManager
               )

set.seed(4)
trainIndices = sample(seq(1:nrow(Attrition)), round(0.7 * nrow(Attrition)))
trainAttrition = Attrition[trainIndices, ]
testAttrition = Attrition[-trainIndices, ]

# Build the model
model = naiveBayes(trainAttrition[, cols_to_use], trainAttrition$Attrition)
predictions = predict(model, testAttrition[, cols_to_use])
CM = confusionMatrix(table(predictions, testAttrition$Attrition))
CM

iterations = 100
masterAcc = matrix(nrow = iterations)
masterSen = matrix(nrow = iterations)
masterSpec = matrix(nrow = iterations)

for (i in 1:iterations) {
  
  trainIndices = sample(seq(1:length(Attrition$Age)), round(0.7 * length(Attrition$Age)))
  trainAttrition = Attrition[trainIndices, ]
  testAttrition = Attrition[-trainIndices, ]
  
  model = naiveBayes(trainAttrition[, cols_to_use], trainAttrition$Attrition)
  predictions = predict(model, testAttrition[, cols_to_use])
  CM = confusionMatrix(table(predictions, testAttrition$Attrition))
  
  masterAcc[i] = CM$overall["Accuracy"]
  masterSen[i] = CM$byClass["Sensitivity"]
  masterSpec[i] = CM$byClass["Specificity"]
}

# Calculate and print the results
MeanAcc = colMeans(masterAcc)
MeanSen = colMeans(masterSen)
MeanSpec = colMeans(masterSpec)

cat("Average Accuracy:", MeanAcc, "\n")
cat("Average Sensitivity:", MeanSen, "\n")
cat("Average Specificity:", MeanSpec, "\n")



# Now lets add in the challenge set 
CompSet = read.csv("/Users/jackkorbitz/School/MSDS_6306/Unit 8/CaseStudy1CompSet No Attrition.csv", header = TRUE)

# Define the columns in the model (excluding those that won't be used)
cols_to_use = c("Age"
               ,"BusinessTravel"
               ,"DistanceFromHome"
               ,"EnvironmentSatisfaction"
               ,"JobInvolvement"
               ,"JobLevel"
               ,"JobRole"
               ,"JobSatisfaction"
               ,"MaritalStatus"
               ,"MonthlyIncome"
               ,"NumCompaniesWorked"
               ,"OverTime"
               ,"RelationshipSatisfaction"
               ,"StockOptionLevel"
               ,"TotalWorkingYears"
               ,"WorkLifeBalance"
               ,"YearsAtCompany"
               ,"YearsInCurrentRole"
               ,"YearsWithCurrManager"
               )

# Train the model on the original dataset
set.seed(4)
trainIndices = sample(seq(1:nrow(Attrition)), round(0.7 * nrow(Attrition)))
trainAttrition = Attrition[trainIndices, ]
# testAttrition = Attrition[-trainIndices, ]  # Not used for prediction

# Build the model
model = naiveBayes(trainAttrition[, cols_to_use], trainAttrition$Attrition)

# Predict on the new dataset
predictions = predict(model, CompSet[, cols_to_use])

# Combine predictions with ID from the new dataset
results = data.frame(ID = CompSet$ID, Predicted_Attrition = predictions)

# View results
print(results)

# If you want to save the results to a CSV file:
write.csv(results, "/Users/jackkorbitz/School/MSDS_6306/Unit 8/Case1PredictionsKorbitz Attrition.csv", row.names = FALSE)
```
## scrapped model:all outputs
### Didnt meet 60% Specificity on 100 itterations
```{r}
# Load data
Attrition = read.csv("/Users/jackkorbitz/School/MSDS_6306/Unit 8/CaseStudy1-data.csv", header = TRUE)

# Define the columns in the model (not shown the amount of combinations tried...)
# Excludes to not be used 1,3,10,11
cols_to_use = c(2  # Age
               ,4  # BusinessTravel
               ,5  # DailyRate
               ,6  # Department
               ,7  # DistanceFromHome
               ,8  # Education
               ,9  # EducationField
               ,12 # EnvironmentSatisfaction
               ,13 # Gender
               ,14 # HourlyRate
               ,15 # JobInvolvement
               ,16 # JobLevel
               ,17 # JobRole
               ,18 # JobSatisfaction
               ,19 # MaritalStatus
               ,20 # MonthlyIncome
               ,21 # MonthlyRate
               ,22 # NumCompaniesWorked
               ,23 # Over18
               ,24 # OverTime
               ,25 # PercentSalaryHike
               ,26 # PerformanceRating
               ,27 # RelationshipSatisfaction
               ,28 # StandardHours    
               ,29 # StockOptionLevel
               ,30 # TotalWorkingYears
               ,31 # TrainingTimesLastYear
               ,32 # WorkLifeBalance
               ,33 # YearsAtCompany
               ,34 # YearsInCurrentRole
               ,35 # YearsSinceLastPromotion
               ,36 # YearsWithCurrManager
               )

set.seed(4)
trainIndices = sample(seq(1:nrow(Attrition)), round(0.7 * nrow(Attrition)))
trainAttrition = Attrition[trainIndices, ]
testAttrition = Attrition[-trainIndices, ]

# Build the model
model = naiveBayes(trainAttrition[, cols_to_use], trainAttrition$Attrition)
predictions = predict(model, testAttrition[, cols_to_use])
CM = confusionMatrix(table(predictions, testAttrition$Attrition))
CM

iterations = 100
masterAcc = matrix(nrow = iterations)
masterSen = matrix(nrow = iterations)
masterSpec = matrix(nrow = iterations)

for (i in 1:iterations) {
  
  trainIndices = sample(seq(1:length(Attrition$Age)), round(0.7 * length(Attrition$Age)))
  trainAttrition = Attrition[trainIndices, ]
  testAttrition = Attrition[-trainIndices, ]
  
  model = naiveBayes(trainAttrition[, cols_to_use], trainAttrition$Attrition)
  predictions = predict(model, testAttrition[, cols_to_use])
  CM = confusionMatrix(table(predictions, testAttrition$Attrition))
  
  masterAcc[i] = CM$overall["Accuracy"]
  masterSen[i] = CM$byClass["Sensitivity"]
  masterSpec[i] = CM$byClass["Specificity"]
}

# Calculate and print the results
MeanAcc = colMeans(masterAcc)
MeanSen = colMeans(masterSen)
MeanSpec = colMeans(masterSpec)

cat("Average Accuracy:", MeanAcc, "\n")
cat("Average Sensitivity:", MeanSen, "\n")
cat("Average Specificity:", MeanSpec, "\n")
```